Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-07-12 paper/paper.tex
paper -> fuer korrektur

\footnotesize
\begin{lstlisting}[language=C++, frame=single, float=*,
	caption=Robot configuration space class structure.,
	label=code4
]
template<int ndof>
class RobotConfigspace : public Configspace<ndof>
{
public:
  RobotConfigspace(const Robot<ndof>* robot_,
                   const polytope *polys_,
                   const int* sys_,
                   const int N_,
                   const int *from_, const int *to_,
                   const int M_,
                   const float* mins_, const float* maxs_, const float dq_,
                   const int nbuf_);
  int init(const int ressource_rank=0, const int ressource_size=1);
  int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
  .
  .
  .
private:
  const Robot<ndof>* robot; //host object
  Robot<ndof>* robotdev;    //GPU object
  collision4::polytope4data* polydata;    //host object
  collision4::polytope4data* polydatadev; //GPU object
  .
  .
  .
}
\end{lstlisting}
\normalsize

@@

struct graph{
    std::map<int,block*> map; //map for accessing blocks
    std::vector<block> blocks;
    std::vector<float> qstorage; //length ndof*N
    int newblockpos;  //position of next new block
    int blocknum;     //number of used blocks
    .
    .
};
\end{lstlisting}
\normalsize

@@

\begin{figure*}
\begin{center}
\begin{tikzpicture}
	\tikzset{mywidth/.style={minimum width=2.5cm},};
	\node[block, mywidth] (l1) {\textbf{Worker 1}};
	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
	\node[frameless, mywidth, below=of l6] (l7) {};
	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
										receive\\ result, \\
										store nodes \\ 
										and edges
									\end{tabular}};
	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};

@@

\end{tikzpicture}
\end{center}
\label{figversion3}
\caption{Structure of asynchronous kernel launches.}
\end{figure*}

@@

\begin{figure}%{r}{0.5\textwidth}
	\label{robotsolution}
	\centering
	\includegraphics[width=0.5\textwidth]{robotsolution2_5.pdf}
	\caption{PRM generated collision-free movement for a 4-axis robot}
\end{figure}

Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
Because of the dependency on random numbers, the algorithm takes different amounts of time for different runs.
Therefore, 10 runs were made for each version and number of processes.
However, the different versions lead to a different average amount of GPU threads needed.
Hence, to be able to better compare how they scale, the time per number of GPU threads is used as a quantity.
Table 1 shows the results.
One can see, that the version with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
The versions with less communication, however, have better absolute times.
Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.

\begin{table}[h]\label{resulttable}
\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
\centering
%\begin{tabular}{| c | c | r | r | r |}

@@

\hline
version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
		& GPUs		& (ms)		& threads		& threads (ms) \\ 
\hline
		& 1			& 1002		& 696413		& 1448	\\ 
1		& 2			& 763		& 838938		& 918	\\ 
		& 4			& 706		& 1113781	& 647	\\ 
\hline
		& 1			& 900		& 608172		& 1490	\\ 
2		& 2			& 647		& 608172		& 1070	\\ 
		& 4			& 568		& 608172		& 947	\\ 
\hline
		& 1			& 696		& 599847		& 1181	\\ 		
3		& 2			& 554		& 599847		& 951	\\ 
		& 4			& 500		& 599847		& 864	\\ 
\hline

@@

As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
In most cases, the GPU code was around 10 times faster. 
However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.

>>>>

\footnotesize
\begin{lstlisting}[language=C++, frame=single, float=*,
	caption=Robot configuration space class structure.,
	label=code4
]
template<int ndof>
class RobotConfigspace : public Configspace<ndof>
{
public:
  RobotConfigspace(const Robot<ndof>* robot_,
                   const polytope *polys_,
                   const int* sys_,
                   const int N_,
                   const int *from_, const int *to_,
                   const int M_,
                   const float* mins_, const float* maxs_, const float dq_,
                   const int nbuf_);
  int init(const int ressource_rank=0, const int ressource_size=1);
  int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
  .
  .
  .
private:
  const Robot<ndof>* robot; //host object
  Robot<ndof>* robotdev;    //GPU object
  collision4::polytope4data* polydata;    //host object
  collision4::polytope4data* polydatadev; //GPU object
  collision4::polytope4data_restrict* polydatadev_restrict; //restricted pointers collection
  .
  .
  .
}
\end{lstlisting}
\normalsize

@@

struct graph{
    std::map<int,block*> map; //map for accessing blocks
    std::vector<block> blocks;
    std::vector<float> qstorage; //length ndof*N
    int newblockpos;  //position of next new block
    int blocknum;     //number of used blocks
    .
    .
    .
};
\end{lstlisting}
\normalsize

@@

\begin{figure}
\begin{center}
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}
	\tikzset{mywidth/.style={minimum width=2.5cm},};
	\node[block, mywidth] (l1) {\textbf{Worker 1}};
	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
										generate\\ nodes, \\
										start kernel
									\end{tabular}};
	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
	\node[frameless, mywidth, below=of l6] (l7) {};
	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
										receive\\ result, \\
										store nodes \\ 
										and edges
									\end{tabular}};
	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};

@@

\end{tikzpicture}
}
\end{center}
\label{figversion3}
\caption{Structure of asynchronous kernel launches.}
\end{figure}

@@

\begin{figure}%{r}{0.5\textwidth}
	\centering
	\includegraphics[width=0.5\textwidth]{robotsolution2_5.pdf}
	\caption{PRM generated collision-free movement for a 4-axis robot}\label{robotpic}
\end{figure}

Tests have been made for a scenario with a robot with 4 joints, shown in figure \ref{robotpic}.
This scenario is difficult because of the small path to be passed in the middle.
For creating and plotting the environment and the resulting trajectories, Matlab functions were written.
Because of the dependency on random numbers, the algorithm takes different amounts of time for different runs.
Therefore, 10 runs were made for each version and number of processes.
However, the different versions lead to a different average amount of GPU threads needed.
Hence, to be able to better compare how they scale, the time per number of GPU threads is used as a quantity.
Table 1 shows the results.

%\begin{table}[h]\label{resulttable}
%\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
%\centering
%%\begin{tabular}{| c | c | r | r | r |}
%
%\begin{tabular}{ c  c  r  r  r }
%
%\hline
%version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
%		& GPUs		& (ms)		& threads		& threads (ms) \\ 
%\hline
%		& 1			& 1002		& 696413		& 1448	\\ 
%1		& 2			& 763		& 838938		& 918	\\ 
%		& 4			& 706		& 1113781	& 647	\\ 
%\hline
%		& 1			& 900		& 608172		& 1490	\\ 
%2		& 2			& 647		& 608172		& 1070	\\ 
%		& 4			& 568		& 608172		& 947	\\ 
%\hline
%		& 1			& 696		& 599847		& 1181	\\ 
%3		& 2			& 554		& 599847		& 951	\\ 
%		& 4			& 500		& 599847		& 864	\\ 
%\hline
%
%\end{tabular}
%\end{table}

\begin{table}[h]\label{resulttable}
\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Eight-Core CPU E5-2650 0 @ 2.00GHz and 4x Tesla K20m.}
\centering
%\begin{tabular}{| c | c | r | r | r |}

@@

\hline
version	&  processes/	& time		& GPU 		& time per $10^6$  \\ 
		& GPUs		& (ms)		& threads		& threads (ms) \\ 
\hline
		& 1			& 6864		& 6171525	& 1070	\\ 
1		& 2			& 1580		& 4489213	& 342	\\ 
		& 4			& 869		& 5853763	& 149	\\ 
\hline
		& 1			& 2136		& 6933768	& 337	\\ 
2		& 2			& 1778		& 6933768	& 290	\\ 
		& 4			& 1677		& 6933768	& 279	\\ 
\hline
		& 1			& 2492		& 19096275	& 130	\\ 
3		& 2			& 1345		& 19096275	& 72		\\ 
		& 4			& 942		& 19096275	& 52		\\ 
\hline
3		& 4			& 32730		& 10038198	& 3281	\\ 
 (CPU	& 8			& 17676		& 10038198	& 1785	\\ 
only)		& 16			& 9936		& 10038198	& 1012	\\ 
\hline

@@

As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
The timing results for the last one are shown in the last section of table 1.
One can see an improvement of up to a factor around 10.
However, the robot code is not optimized for the CPU version, as the GPU kernel is just rewritten with an for loop.
Therefore this comparison is not quite fair.
